---
title: "Readiness to Change"
author: "Benjamin Carter"
date: "6/20/2020"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 3
    toc_float: yes
  slidy_presentation:
    incremental: yes
  ioslides_presentation:
    incremental: yes
    widescreen: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
library(dplyr)
library(gmodels)
library(ggplot2)
# paths V:\Depts\CtrTransRsch\Dept Private\CTR Staff\PROJECT - Readiness to change manuscript\DATA
data.dir.path <- file.path(
  "V:",
  "Depts", 
  "CtrTransRsch", 
  "Dept Private", 
  "CTR Staff", 
  "PROJECT - Readiness to change manuscript", 
  "DATA"
  )
```

```{r loadData, include=FALSE}
df <- readxl::read_xlsx(
  file.path(data.dir.path, "Nick_Rural16_OAT 6-11-20.xlsx")
  )

```


```{r selection, include=FALSE}
df.processed <- df %>% 
  mutate(
    Gender = ifelse(
      Gender == 1,
      "Female",
      "Male"
      ),
    rtc_diet = ifelse(
      `1a_pre` == 1,
      "precontemplative",
      ifelse(
        `1a_pre` == 2,
        "contemplative",
        ifelse(
          `1a_pre` == 3,
          "preparation",
          ifelse(
            `1a_pre` == 4,
            "action",
            ifelse(
              `1a_pre` == 5,
              "maintain",
              `1a_pre`
            )
          )
        )
      )
    ),
    rtc_exercise = ifelse(
      `2a_pre` == 1,
      "precontemplative",
      ifelse(
        `2a_pre` == 2,
        "contemplative",
        ifelse(
          `2a_pre` == 3,
          "preparation",
          ifelse(
            `2a_pre` == 4,
            "action",
            ifelse(
              `2a_pre` == 5,
              "maintain",
              `2a_pre`
            )
          )
        )
      )
    ),
    rtc_diet_groups = ifelse(
      `1a_pre` == 1,
      "Pre/Contemplative",
      ifelse(
        `1a_pre` == 2,
        "Pre/Contemplative",
        ifelse(
          `1a_pre` == 3,
          "Preparation",
          ifelse(
            `1a_pre` == 4,
            "Action",
            ifelse(
              `1a_pre` == 5,
              "Maintain",
              `1a_pre`
            )
          )
        )
      )
    ),
    rtc_exercise_groups = ifelse(
    `2a_pre` == 1,
    "Pre/Contemplative",
    ifelse(
      `2a_pre` == 2,
      "Pre/Contemplative",
      ifelse(
        `2a_pre` == 3,
        "Preparation",
        ifelse(
          `2a_pre` == 4,
          "Action",
          ifelse(
            `2a_pre` == 5,
            "Maintain",
            `2a_pre`
          )
        )
      )
    )
  ),
  Include = ifelse(
    is.na(`1a_pre`) |
      is.na(`2a_pre`),
    "FALSE",
    "TRUE"
  ),
  MaxWeightLost = MaxWeightLost*-1
) %>% 
  select(
    rtc_diet,
    rtc_diet_groups,
    rtc_exercise,
    rtc_exercise_groups,
    Include,
    Gender,
    Age,
    `Start weight` = Todays_Weight,
    BMI,
    `Maximum weight lost` = MaxWeightLost,
    `Percent weight lost` = pWeightLost,
    `Lost 7 percent during core` = Met7GoalCore,
    `Lost 7 percent overall` = Met7GoalOVL,
    `Lost 5 percent during core` = Met5GoalCore,
    `Lost 5 percent overall` = Met5GoalOVL,
    `Core attendance` = Core,
    `Post-core attendance` = AfterCore,
    `Overall attendance` = `Number of Sessions`,
    `Met fat goal` = FatGramGoalMet,
    `Met fat goal (percentage of sessions)` = Met_Fat_Gram,
    `Met exercise goal at least 50 percent` = Exercise_Goal_50_Met,
    `Met exercise goal (percentage of sessions)` = Met_Activity_Goal
  ) %>% 
  mutate(
    rtc_diet_groups = as.factor(
      rtc_diet_groups
      ),
    rtc_exercise_groups = as.factor(
      rtc_exercise_groups
      ),
    `Lost 7 percent during core` = ifelse(
      `Lost 7 percent during core` == 1,
      TRUE,
      FALSE
      ),
    `Lost 7 percent overall` = ifelse(
      `Lost 7 percent overall` == 1,
      TRUE,
      FALSE
      ),
    `Lost 5 percent during core` = ifelse(
      `Lost 5 percent during core` == 1,
      TRUE,
      FALSE
    ),
    `Lost 5 percent overall` = ifelse(
      `Lost 5 percent overall` == 1,
      TRUE,
      FALSE
    ),
    `Met fat goal` = ifelse(
      `Met fat goal` == 1,
      TRUE,
      FALSE
    ),
    `Met exercise goal at least 50 percent` = ifelse(
      `Met exercise goal at least 50 percent` == 1,
      TRUE,
      FALSE
    )
  )

df.processed$rtc_diet_groups <- factor(
  df.processed$rtc_diet_groups,
  levels = c("Pre/Contemplative", "Preparation", "Action", "Maintain")
  )

df.processed$rtc_exercise_groups <- factor(
  df.processed$rtc_exercise_groups,
  levels = c("Pre/Contemplative", "Preparation", "Action", "Maintain")
  )

```


# Descriptives

## Comparison between included/excluded individuals

```{r message=FALSE, warning=FALSE}
tab1 <- compareGroups::compareGroups(
  Include ~ .- rtc_diet_groups - rtc_exercise - rtc_exercise_groups - rtc_diet,
  df.processed
  )

compareGroups::createTable(
  tab1,
  show.p.overall = FALSE
  ) %>% 
  compareGroups::export2md()

```

## Summary statistics of included participants

```{r}
df.processed <- df.processed %>% 
  filter(
    Include == "TRUE"
  )

tab1 <- compareGroups::compareGroups(
   ~ . - Include - rtc_diet_groups - rtc_exercise - rtc_exercise_groups - rtc_diet,
  df.processed
  )

compareGroups::createTable(
  tab1,
  show.n = FALSE
  ) %>% 
  compareGroups::export2md()
```


# Comparison of categorical variables by readiness to change

Categorical variables (i.e. variables that were nominal rather than quantitative in nature) were examined in relation to readiness to change diet or exercise using a Pearson's chi-square analysis. What follows are the reported $\chi^2$ statistics, degrees of freedom, and p-values from each test.

## Readiness to change diet

```{r categoricals, message=FALSE,warning=FALSE,echo=FALSE}

binary_vars <- list(
  "Lost 7 percent during core" = "Lost 7 percent during core",
  "Lost 7 percent overall" = "Lost 7 percent overall",
  "Lost 5 percent during core" = "Lost 5 percent during core",
  "Lost 5 percent overall" = "Lost 5 percent overall",
  "Met fat goal" = "Met fat goal",
  "Met exercise goal at least 50 percent" = "Met exercise goal at least 50 percent"
  )

chi_diet <- function(x){
  chisq <- chisq.test(
    df.processed$rtc_diet_groups,
    df.processed[[x]],
    correct = TRUE
  )
  return(chisq)
}

chis.diet <- lapply(binary_vars, chi_diet)

df.stats <- data.frame(
  "Variable"=character(),
  "X-squared"=numeric(),
  "df"=integer(),
  "p-value"=integer(),
  row.names = NULL
)

for (i in names(chis.diet)){
  row.to.add <- data.frame(
    "Variable" = i,
    "X-squared" = chis.diet[[i]]$statistic,
    "df" = chis.diet[[i]]$parameter,
    "p-value" = chis.diet[[i]]$p.value,
    row.names = NULL
    )
  df.stats <- rbind(df.stats, row.to.add)
}

knitr::kable(
  df.stats,
  output = "pandoc"
) %>% 
  kableExtra::kable_styling(
    bootstrap_options = c(
      "striped", 
      "hover", 
      "responsive"
      )
    )

```

## Readiness to change exercise

```{r}

chi_exercise <- function(x){
  chisq <- chisq.test(
    df.processed$rtc_exercise_groups,
    df.processed[[x]],
    correct = TRUE
  )
  return(chisq)
}

chis.exercise <- lapply(binary_vars, chi_exercise)

df.stats <- data.frame(
  "Variable"=character(),
  "X-squared"=numeric(),
  "df"=integer(),
  "p-value"=integer(),
  row.names = NULL
)

for (i in names(chis.exercise)){
  row.to.add <- data.frame(
    "Variable" = i,
    "X-squared" = chis.exercise[[i]]$statistic,
    "df" = chis.exercise[[i]]$parameter,
    "p-value" = chis.exercise[[i]]$p.value,
    row.names = NULL
    )
  df.stats <- rbind(df.stats, row.to.add)
}

knitr::kable(
  df.stats,
  output = "pandoc"
) %>% 
  kableExtra::kable_styling(
    bootstrap_options = c(
      "striped", 
      "hover", 
      "responsive"
      )
    )

```

# Comparison of continuous variables by readiness to change

The effect of readiness to change on quantitative measures were examined using via an analysis of variance. The appropriate statistics from each test are found below.

## Readiness to change diet

```{r continuous, message=FALSE}

cont_vars <- list(
  "Core attendance" = 
    "Core attendance",
  "Post-core attendance" = 
    "Post-core attendance",
  "Overall attendance" = 
    "Overall attendance",
  "Maximum weight lost" = 
    "Maximum weight lost",
  "Percent weight lost" = 
    "Percent weight lost",
  "Met fat goal (percentage of sessions)" = 
    "Met fat goal (percentage of sessions)",
  "Met exercise goal (percentage of sessions)" = 
    "Met exercise goal (percentage of sessions)"
)

aov_readiness <- function(x, y){
  aov.results <- aov(
    df.processed[[x]] ~ df.processed[[y]],
    data = df.processed
      )
  
  return(aov.results)
}

aov_diet <- lapply(cont_vars, aov_readiness, y = "rtc_diet_groups")

#lapply(aov_diet, summary)

df.stats <- data.frame(
  "Variable"=character(),
  "F-statistic"=numeric(),
  "df"=integer(),
  "p-value"=integer(),
  row.names = NULL
)

for (i in names(aov_diet)){
  a <- summary(aov_diet[[i]])
  row.to.add <- data.frame(
    "Variable" = i,
    "F-statistic" = a[[1]]$`F value`[1],
    "df" = a[[1]]$`Df`[1],
    "p-value" = a[[1]]$`Pr(>F)`[1],
    row.names = NULL
    )
  df.stats <- rbind(df.stats, row.to.add)
}

knitr::kable(
  df.stats,
  output = "pandoc"
) %>% 
  kableExtra::kable_styling(
    bootstrap_options = c(
      "striped", 
      "hover", 
      "responsive"
      )
    )

```

## Readiness to change exercise

```{r continuous2}

aov_exercise <- lapply(cont_vars, aov_readiness, y = "rtc_exercise_groups")

# lapply(aov_exercise, summary)

df.stats <- data.frame(
  "Variable"=character(),
  "F-statistic"=numeric(),
  "df"=integer(),
  "p-value"=integer(),
  row.names = NULL
)

for (i in names(aov_exercise)){
  a <- summary(aov_exercise[[i]])
  row.to.add <- data.frame(
    "Variable" = i,
    "F-statistic" = a[[1]]$`F value`[1],
    "df" = a[[1]]$`Df`[1],
    "p-value" = a[[1]]$`Pr(>F)`[1],
    row.names = NULL
    )
  df.stats <- rbind(df.stats, row.to.add)
}

knitr::kable(
  df.stats,
  output = "pandoc"
) %>% 
  kableExtra::kable_styling(
    bootstrap_options = c(
      "striped", 
      "hover", 
      "responsive"
      )
    )

```

# Interactions in readiness to change

```{r interactions}

cross.both <- chisq.test(
  df.processed$rtc_diet_groups,
  df.processed$rtc_exercise_groups,
  correct = TRUE
  )

cv.test = function(x,y) {
  CV = sqrt(chisq.test(x, y, correct=FALSE)$statistic /
    (length(x) * (min(length(unique(x)),length(unique(y))) - 1)))
  return(as.numeric(CV))
}

CV <- cv.test(
  df.processed$rtc_diet_groups,
  df.processed$rtc_exercise_groups
  )

# fmla <-  as.formula("df.processed$`Maximum weight lost` ~ df.processed$rtc_diet_groups*df.processed$rtc_exercise_groups, df.processed")
```

Responses to the readiness to change questions are probably related correlated. This would make sense as individuals who are willing to change their diet are probably more likely to change their exercise as well. The converse is probably also true. Cramer's V between both questions is `r CV`. This interaction is incorporated into the predictive models.

```{r}
knitr::kable(
  as.data.frame(cross.both$observed) %>% tidyr::spread(key = Var2, value = Freq)) %>% 
  kableExtra::kable_styling(
    bootstrap_options = c(
      "striped", 
      "hover", 
      "responsive"
      )
    )

```

# Predictive models for success

Statistical models were created in an attempt to predict outcomes based on patient demographics and baseline response to the readiness to change survey questions. The following measures were entered into the model: starting weight, age, gender, BMI, readiness to change diet, readiness to change exercise. An interaction between readiness to change diet and readiness to change exercise was included in the model. A logistic model was used to predict binary outcomes (e.g. did they lose 7% body weight by the end of the study). A linear model was used to predict countinuous outcomes (e.g. what percentage of their body weight were they able to lose).

## Logistic models for categorical outcomes

```{r}

model_log <- function(x){

  model <- glm(
    formula = 
      df.processed[[x]] ~ 
      rtc_diet_groups + 
      rtc_exercise_groups, 
    data = df.processed, 
    family = binomial
    ) %>% 
    summary()
  
  return(
    as.data.frame(
      model$coefficients
      )
    )
}

binary_list <- lapply(binary_vars, model_log)

for (i in names(binary_list)){
  
  b <- binary_list[[i]]$`Pr(>|z|)`
  
  sig.p <- ifelse(
    0.05 >=
    min(
      b[2:length(b)]
    ),
    TRUE,
    FALSE
  )
  
  if (sig.p == TRUE) {
    print("==============================================")
    print(i)
    print(binary_list[[i]][c(1,4)])
  }  
}

```

## Linear model for continuous outcomes

```{r}

model_lm <- function(x){
  
  model <- lm(
    formula = 
      df.processed[[x]] ~ 
      rtc_diet_groups + 
      rtc_exercise_groups, 
    data = df.processed
    ) %>% 
    summary()
  
  a <- data.frame(
        "Beta estimate" = model$coefficients[,1],
        "p-value" = model$coefficients[,4]
        )
  
  return(a)
}

continuous_list <- lapply(cont_vars, model_lm)

for (i in names(continuous_list)){
  
  b <- continuous_list[[i]]$p.value
  
  sig.p <- ifelse(
    0.05 >=
    min(
      b[2:length(b)]
    ),
    TRUE,
    FALSE
  )
  
  if (sig.p == TRUE) {
    print("===============================================")
    print(i)
    print(continuous_list[[i]])
  }  
}

```




# Core Completers

Did completing the 16 core sessions predict anything?

```{r}
df.processed <- df.processed %>% 
  mutate(
    core_completion = ifelse(
      `Core attendance` >= 16,
      "Complete",
      "Incomplete"
    )
  )

comp_tab <- compareGroups::compareGroups(core_completion ~ .- rtc_diet_groups - rtc_exercise - rtc_exercise_groups - rtc_diet - Include - `Core attendance`,
                             data = df.processed)

compareGroups::createTable(
  comp_tab,
  show.p.overall = FALSE
) %>% 
  compareGroups::export2md(
    caption = "A comparison of participants according to whether they completed the 16 core sessions."
    )

```

## Categorical variables

Categorical variables (i.e. variables that were nominal rather than quantitative in nature) were examined in relation to whether they completed the 16 core sessions using a Pearson's chi-square analysis. What follows are the reported $\chi^2$ statistics, degrees of freedom, and p-values from each test.

```{r}
binary_vars <- list(
  "Lost 7 percent during core" = "Lost 7 percent during core",
  "Lost 7 percent overall" = "Lost 7 percent overall",
  "Lost 5 percent during core" = "Lost 5 percent during core",
  "Lost 5 percent overall" = "Lost 5 percent overall",
  "Met fat goal" = "Met fat goal",
  "Met exercise goal at least 50 percent" = "Met exercise goal at least 50 percent"
  )

chi_diet <- function(x){
  chisq <- chisq.test(
    df.processed$core_completion,
    df.processed[[x]],
    correct = TRUE
  )
  return(chisq)
}

chis.diet <- lapply(binary_vars, chi_diet)

df.stats <- data.frame(
  "Variable"=character(),
  "X-squared"=numeric(),
  "df"=integer(),
  "p-value"=integer(),
  row.names = NULL
)

for (i in names(chis.diet)){
  row.to.add <- data.frame(
    "Variable" = i,
    "X-squared" = chis.diet[[i]]$statistic,
    "df" = chis.diet[[i]]$parameter,
    "p-value" = chis.diet[[i]]$p.value,
    row.names = NULL
    )
  df.stats <- rbind(df.stats, row.to.add)
}

knitr::kable(
  df.stats,
  output = "pandoc"
) %>% 
  kableExtra::kable_styling(
    bootstrap_options = c(
      "striped", 
      "hover", 
      "responsive"
      )
    )

# graphs

df.processed %>% 
  tidyr::gather(
    key = Variables,
    value = Value,
    c(
      "Lost 7 percent during core",
      "Lost 7 percent overall",
      "Lost 5 percent during core",
      "Lost 5 percent overall",
      "Met fat goal",
      "Met exercise goal at least 50 percent"
      )
  ) %>% 
  ggplot(
    aes(core_completion)
  ) +
  geom_bar(
    aes(fill = Value),
    position = "fill",
    alpha = 0.8
  ) +
  facet_wrap("Variables") +
  theme_classic() +
  labs(
    title = "Proportion of Patients Achieving Program Goals",
    x = "Core Completion Status",
    y = "Proportion",
    fill = "Did They Meet Goal?"
  )


```

## Continuous variables

The effect of completing the core sessions on quantitative measures were examined using via an analysis of variance. The appropriate statistics from each test are found below.

```{r}

cont_vars <- list(
  "Post-core attendance" = 
    "Post-core attendance",
  "Overall attendance" = 
    "Overall attendance",
  "Maximum weight lost" = 
    "Maximum weight lost",
  "Percent weight lost" = 
    "Percent weight lost",
  "Met fat goal (percentage of sessions)" = 
    "Met fat goal (percentage of sessions)",
  "Met exercise goal (percentage of sessions)" = 
    "Met exercise goal (percentage of sessions)"
)

aov_readiness <- function(x, y){
  aov.results <- aov(
    df.processed[[x]] ~ df.processed[[y]],
    data = df.processed
      )
  
  return(aov.results)
}

aov_diet <- lapply(cont_vars, aov_readiness, y = "rtc_diet_groups")

# lapply(aov_diet, summary)

df.stats <- data.frame(
  "Variable"=character(),
  "F-statistic"=numeric(),
  "df"=integer(),
  "p-value"=integer(),
  row.names = NULL
)

for (i in names(aov_diet)){
  a <- summary(aov_diet[[i]])
  row.to.add <- data.frame(
    "Variable" = i,
    "F-statistic" = a[[1]]$`F value`[1],
    "df" = a[[1]]$`Df`[1],
    "p-value" = a[[1]]$`Pr(>F)`[1],
    row.names = NULL
    )
  df.stats <- rbind(df.stats, row.to.add)
}

knitr::kable(
  df.stats,
  output = "pandoc"
) %>% 
  kableExtra::kable_styling(
    bootstrap_options = c(
      "striped", 
      "hover", 
      "responsive"
      )
    )

df.processed %>% 
  tidyr::gather(
    key = Variables,
    value = Value,
    c(
      "Post-core attendance",
      "Overall attendance",
      "Maximum weight lost",
      "Percent weight lost",
      "Met fat goal (percentage of sessions)",
      "Met exercise goal (percentage of sessions)"
      )
  ) %>% 
  ggplot(
    aes(core_completion, Value)
  ) +
  geom_boxplot() +
  facet_wrap(
    "Variables",
    scales = "free"
    ) +
  theme_classic() +
  labs(
    title = "Continuous Outcomes for Completers v. Non-Completers",
    x = "Group",
    y = ""
  )

```

```{r}
ggplot(
  df.processed,
  aes(Age)
) +
  geom_boxplot()
```


